Why Real-World Subgraph Isomorphism is Easy, \\ and What This Means for Graph Databases

### 2016-09-28 paper/paper.tex
More data

The \emph{non-induced subgraph isomorphism problem} is to find an injective mapping from a given
pattern graph to a given target graph which preserves adjacency---in essence, we are ``finding a
copy of'' the pattern inside the target. The \emph{induced} variant of the problem additionally
requires that the mapping preserve non-adjacency, so there are no ``extra edges'' in the copy of the
pattern that we find. We illustrate both variants in \cref{figure:sip}.  Although these problems
are \NP-complete, modern practical subgraph isomorphism algorithms can handle problem instances
with many hundreds of vertices in the pattern graph, and up to ten thousand vertices in the target
graph \cite{Cordella:2004,Solnon:2010,Audemard:2014,McCreesh:2015}, leading to successful
application in areas such as computer vision \cite{Damiand:2011,Solnon:2015}, biochemistry
\cite{Giugno:2013,Carletti:2015}, and pattern recognition \citep{Conte:2004}.  In particular,
subgraph isomorphism is one of the key components in supporting complex queries in graph
databases---typically, these systems store a fixed set of target graphs, and for a sequence of
pattern queries, they must return every target graph which contains that pattern.  One common
approach to this problem is to combine a subgraph isomorphism algorithm with an indexing system in a
so-called ``filter / verify'' model, where invariants are used to attempt to pre-exclude
unsatisfiable instances to avoid the cost of a subgraph isomorphism call. In this context, the terms
\emph{matching} and \emph{verification} are often used for the subgraph isomorphism step (or a
slightly broader problem, for example permitting wildcards).

@@

What goes on here?

Min unsat core.

Insufficient vertices.

VF2 does not detect the case that there are insufficient labels. Also empty pattern but
not empty target.

@@

@@

?? Additionally, this stuff works on satisfiable instances too.

@@

We build up a simple algorithm. Start with VF2, adapt to track domains, smallest domain first, all
different. Show its effectiveness at eliminating unsatisfiable instances, and give its total
performance.  None of the steps in this algorithm are new. Our aim in this section is to show which
aspects of modern algorithm design are critical, and that an indexless sequential scan is entirely
plausible with a decent algorithm.

@@

@@

?? They need to read outside the field. AI has a lot to offer. There needs to be much more focus on
the matching algorithm, and how it interacts with the system as a whole.

?? We need optimisations for trivial cases. Lazy forward checking, cheap alternatives to heuristics and
all-different. Presolving. Huge domains. Ongoing use of VF2 and micro vs millisecond benchmarks.

?? Degrade to clique.

>>>>

The \emph{non-induced subgraph isomorphism problem} is to find an injective mapping from a given
pattern graph to a given target graph which preserves adjacency---in essence, we are ``finding a
copy of'' the pattern inside the target. The \emph{induced} variant of the problem additionally
requires that the mapping preserve non-adjacency, so there are no ``extra edges'' in the copy of the
pattern that we find. We illustrate both variants in \cref{figure:sip}.  Although these problems
are \NP-complete, modern practical subgraph isomorphism algorithms can handle problem instances
with many hundreds of vertices in the pattern graph, and up to ten thousand vertices in the target
graph \cite{Cordella:2004,Solnon:2010,Audemard:2014,McCreesh:2015}, leading to successful
application in areas such as computer vision \cite{Damiand:2011,Solnon:2015}, biochemistry
\cite{Giugno:2013,Carletti:2015}, and pattern recognition \citep{Conte:2004}.  In particular,
a labelled version of subgraph isomorphism is one of the key components in supporting complex
queries in graph databases---typically, these systems store a fixed set of target graphs, and for a
sequence of pattern queries, they must return every target graph which contains that pattern.  One
common approach to this problem is to combine a subgraph isomorphism algorithm with an indexing
system in a so-called ``filter / verify'' model, where invariants are used to attempt to pre-exclude
unsatisfiable instances to avoid the cost of a subgraph isomorphism call. In this context, the terms
\emph{matching} and \emph{verification} are often used for the subgraph isomorphism step (or a
slightly broader problem, for example permitting wildcards).

@@

?? What goes on here? With labels, there are two regions where VF2 often finds instances really
hard, but Glasgow and LAD do not. Looking at the data more closely, although the average runtime is
high, there is a wide distribution of runtimes in these areas, and VF2 does sometimes find these
instances trivial too.

?? Give examples here: VF2 does not detect simple things like there being two red vertices in the
pattern but one in the target, or even no red vertices in the target. In contrast, Glasgow and LAD
do. Explain thrashing.

?? Explain what's going on in the satisfiable region. This is a bit more subtle: is it to do with
neighbours, rather than a root node property? Would be nice if we could find a representative
example.

?? Look more at size of min unsat core in these areas?

@@

?? Give a quick example here of how an index would eliminate the two red vertices in the pattern but
one red vertex in the target example.

@@

?? Additionally, this stuff works on satisfiable instances too. Indexing addresses some of the
obviously-unsatisfiable cases that VF2 struggles with, but does nothing on the satisfiable
instances.

@@

?? We build up a simple algorithm. Use smallest domain first, all different, NDS, possibly distance
2 filtering. Show its effectiveness at eliminating unsatisfiable instances from graph database test
sets, and give its total performance.  None of the steps in this algorithm are new. Our aim in this
section is to show which aspects of modern algorithm design are critical, and that an indexless
sequential scan is entirely plausible with a decent algorithm.

?? To tie this into earlier parts of the paper, we also show what happens if we use the VF2
heuristic in this setting, rather than smallest domains first.

?? This stuff eliminates all the spuriously hard unsatisfiable instances along the bottom of the
graph, but also helps on the satisfiable instances that an indexing system cannot touch.

@@

?? Clearly lots of scope for cooperation between AI and graph databases. AI is not just good for
better algorithms, but the design of entire systems.

@@

?? We need optimisations for trivial cases. Lazy forward checking, cheap alternatives to heuristics and
all-different. Presolving. Huge domains. Ongoing use of VF2 and micro vs millisecond benchmarks.
Precalculating NDS, distance 2 etc, not for indexing. Edge labels.

?? Degrade to clique. Symmetries.
